---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
setwd("~/Desktop/R/ass/")
library(data.table)
library(lubridate)
library(ggplot2)
activity <- fread("activity.csv")
activity[, date:=ymd(date, tz = "EST")]
```


## What is mean total number of steps taken per day?
```{r}
daily <- activity[, .(daily=sum(steps, na.rm = T)), by=date]
ggplot(data=daily, aes(x=daily)) + geom_histogram(binwidth=1000)
mean.daily <- round(mean(daily$daily))
median.daily <- round(median(daily$daily))
```
The mean of the total number of steps taken per day is `r mean.daily`. 
The median of the total number of steps taken per day is `r median.daily`.


## What is the average daily activity pattern?
```{r}
activity[, ave:=mean(steps, na.rm = T), by=interval]
ggplot(data=activity, aes(x=interval, y=ave)) + geom_line()
max.steps <- activity[which.max(ave), interval]
```
The interval with the max number of steps on average across all the days is `r max.steps`.


## Imputing missing values
```{r}
missing.vector <- is.na(activity$steps)
missing <- sum(is.na(activity$steps))
```
The total number of missing values in the dataset is `r missing`.
```{r}
#impute the missing data with averaged number steps taken during that interval of all of the days
activity[missing.vector, steps:=round(ave,0)]
activity.no.na <- activity[,1:3]
daily.no.na <- activity.no.na[, .(daily=sum(steps)), by=date]
ggplot(data=daily.no.na, aes(x=daily)) + geom_histogram(binwidth=1000)
mean.daily.no.na <- round(mean(daily.no.na$daily))
median.daily.no.na <- round(median(daily.no.na$daily))
```
The mean of the total number of steps taken per day is `r mean.daily.no.na`. 
The median of the total number of steps taken per day is `r median.daily.no.na`.
These values are slightly bigger than the estimates from the first part of the assignment. Imputing missing data resulted in slightly higher estimates of the total daily number of steps.


## Are there differences in activity patterns between weekdays and weekends?
```{r}
activity[, Day:=weekdays(date)]
activity[, category:="weekday"]
activity[grep(pattern="S", x=Day), category:="weekend"]
ave.5.by.weekday <- activity[, .(ave=mean(steps)), by=.(interval, category)]
ggplot(data=ave.5.by.weekday, aes(x=interval, y=ave)) + geom_line() + facet_grid(category~.)
```

